---
title: "Evaluation of simulation FACILITATION and COMPETITION - testing parameters "
output: html_notebook
---


Script to try parameters out and explore the output.

I haven't found empirical network data for facilitation and competition so far. The evaluation I am doing consist mainly in exploring the species richness curve (expecting a progressive increase with time and then an estabilization), degree distribution (expecting diversity of values), and some model objects (trait evolution, speciation/extinction events).


```{r}
library(EnvStats)
library(foreach)

source("../functions/functions_simulation_fac_comp.R")

```


For plotting:

```{r}
my.theme<-theme(axis.text=element_text(size=12),
                axis.title = element_text(size = 14),
                legend.position = "top",
                legend.text=element_text(size=10),
                legend.title = element_text(size=12),
                plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
                axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)),
                axis.title.x = element_text(margin = margin(t = 15, r = 0, b = 0, l = 0)))
```


# Species richness





```{r}

plot_spp_richness <- function(steps, pres_mat) {
  
  timesteps <- 1:steps
  
  n_spp_time <- c()
  
  for (i in 1:nrow(pres_mat)) {
  
  n_spp_time[i] <- length(which(pres_mat[i,] == 1))
  
  }
  
  
  df_spprich_time <- data.frame(timesteps,n_spp_time)
  
  p.spp <- ggplot(df_spprich_time, aes(x=timesteps, y = n_spp_time)) +
  geom_line(color="black", linetype="twodash") +
  theme_bw()+
  my.theme+
  xlab("timesteps")+
  ylab("N species")
  
  return(p.spp)
}



```




# set general parameters

```{r}

pars = list()

# General parameters
#pars$int = 0 # 0 for competition, 1 for facilitation, 2 for predation
pars$Smax = 1000 # Maximal number of species in the system

# Related to interaction range trait
pars$av_r = 0.1 # Half range of the niche of the first species
pars$sd = 0.5*pars$av_r + 0.0001 # Standard deviation of the normal distribution used to calculate the niche optimum trait of a new species

pars$u_max = 0.15 # Speciation probability

# Extinction probability, negative interactions communities
pars$a_eneg = 0.025 # Shape of the exponential decay of the negative extinction - interaction relationship
pars$e_0neg = 0.5 # Asymptotic extinction probability with infinite negative interactions
pars$e_1neg = 1  # Extinction probability with absence of interactions

# Extinction probability, positive interactions communities
pars$a_epos = 1.2  # Shape of the exponential decay of the positive extinction - interaction relationship
pars$e_0pos = 0.075 # Asymptotic extinction probability with infinite positive interactions
pars$e_1pos = 5.19 # 1 - pars$e_0pos

# Establishment probability, negative interactions communities
pars$a_uneg = 0.075 # Shape of the exponential decay of the colonization - interaction relationship
pars$u_0neg = 0.075 # Asymptotic establishment probability with infinite competition interactions
pars$u_1neg = 2 # Establishment probability with absence of competition interaction

# Establishment probability, positive interactions communities
pars$a_u = 0.45 # Shape of the exponential decay of the colonization - interaction relationship
pars$u_0 = 1 # Asymptotic establishment probability with infinite facilitation interactions
pars$u_1 = -1 # Establishment probability with absence of facilitation interactions
pars$d = 0.5 # Decrease speed of the establishment probability

# Extinction and establishment probabilitie for both, positive and negative interactions communities
pars$Bspe = 4 # Constant minimal number of interaction per species (extablishment prob)
pars$Bext = 4 # Constant minimal number of interaction per species (extinction prob)
pars$I_max = 40 # Maximal number of interactioning species

```


# set parameters simulation

```{r}
nsteps = 250 # Set the maximum timestep per simulation
nsim = 2 # Set the number of simulations

# Vector to record number of simulation needed to reach 100 valid simulations
needed_sim_pos <- c(rep(NA, nsim))
needed_sim_neg <- c(rep(NA, nsim))
```


# Simulation

choose interaction

```{r}
int = 1
```



```{r}

list_results_fac <- list()
list_results_comp <- list()
seed_list <- NULL

	total_nb_sim <- 0
	b <- 1

	for(int in 0:1){
		pars$int = int # Define the interaction type: 0 = competition ; 1 = facilitation
		successful_sim <- 1 # Set the count of simulations
		list_res <- list() # Set the list to store the results

		if(pars$int == 1){
			print("Positive interctions")
		} else {
			print("Negative interctions")
		}

		seed_record <- c(rep(NA, nsim)) # Set the vector to record the seed that have been used

		while(successful_sim < nsim+1){

  		seed <- sample.int(80000, 1) # Pick a random number, this number will be the seed for one simulation
  		print(seed)

			# If a seed have already been tested, pick another one
			if(length(which(seed_list == seed)) != 0){
				while(seed %in% seed_list){
    			seed <- sample.int(80000, 1)
  			}
			}

			seed_list <- c(seed_list, seed) # Add the seed

			# SIMULATION ; from the file code/functions/functions_M-C_bifurcation.R
  		simulation <- sim_model_bif_fc(seed = seed, pars, nsteps = nsteps)  
  	
  		#---------------------- stored: pres, traits_mat, anc, extinct, Stotact, L, L_list

			total_nb_sim <- total_nb_sim +1 # Count the total amount of simulations

			# Test if we have enough species at the timestep 150 in the simulation
  		success <- sum(simulation$pres[150,])
			#print(paste0("Success = ", success))

  		if(success >= 20){
				#print("the simulation has more than 19 species at the time step 150")
    		seed_record[successful_sim] <- seed # Record seed which alows us to have "good" simulation
    		res_sim <- list(simulation_number = paste0("simulation", successful_sim), 
    		                seed = seed,
    		                parameters = pars, 
    		                presence_matrix = simulation$pres, 
    		                traits_df = simulation$traits,
    		                parentage_matrix = simulation$anc,
    		                extinxction_matrix = simulation$extinct, 
    		                network_list = simulation$L_list, 
    		                anc_dist_table = simulation$dist_anc, 
    		                list_anc_dist = simulation$list_dist_anc) # Record results from the simulation
    		
    		if(int == 1){
    		  list_results_fac[[nsim]] <- res_sim
    		  
    		}else if(int == 0) {
    		  list_results_comp[[nsim]] <- res_sim
    		  
    		}
    		
    		
    		

    		print(paste("simulation ", successful_sim, "sur", nsim))
    		successful_sim <- successful_sim + 1 # Count the amount of "good" simulations
    		
    		
    		
    		
  		}
		}

		
	}


```






# plot species richness over time

```{r}

plot_spp_richness(steps = nsteps, pres_mat = list_results_fac[[2]]$presence_matrix)

```


# Degree distribution - Simulation

