---
title: "From simulations_script.R"
output: html_notebook
---

```{r}
#install.packages("EnvStats")
#library(foreach) #To do parallel computiong

library(EnvStats)
library(foreach)
```


# FACILITATION & COMPETITION




### set general parameters

```{r}
rm(list = ls())

pars = list()

# General parameters
#pars$int = 0 # 0 for competition, 1 for facilitation, 2 for predation
pars$Smax = 1000 # Maximal number of species in the system

# Related to interaction range trait
pars$av_r = 0.1 # Half range of the niche of the first species
pars$sd = 0.5*pars$av_r + 0.0001 # Standard deviation of the normal distribution used to calculate the niche optimum trait of a new species

pars$u_max = 0.15 # Speciation probability

# Extinction probability, negative interactions communities
pars$a_eneg = 0.025 # Shape of the exponential decay of the negative extinction - interaction relationship
pars$e_0neg = 0.5 # Asymptotic extinction probability with infinite negative interactions
pars$e_1neg = 1  # Extinction probability with absence of interactions

# Extinction probability, positive interactions communities
pars$a_epos = 1.2  # Shape of the exponential decay of the positive extinction - interaction relationship
pars$e_0pos = 0.075 # Asymptotic extinction probability with infinite positive interactions
pars$e_1pos = 5.19 # 1 - pars$e_0pos

# Establishment probability, negative interactions communities
pars$a_uneg = 0.075 # Shape of the exponential decay of the colonization - interaction relationship
pars$u_0neg = 0.075 # Asymptotic establishment probability with infinite competition interactions
pars$u_1neg = 2 # Establishment probability with absence of competition interaction

# Establishment probability, positive interactions communities
pars$a_u = 0.45 # Shape of the exponential decay of the colonization - interaction relationship
pars$u_0 = 1 # Asymptotic establishment probability with infinite facilitation interactions
pars$u_1 = -1 # Establishment probability with absence of facilitation interactions
pars$d = 0.5 # Decrease speed of the establishment probability

# Extinction and establishment probabilitie for both, positive and negative interactions communities
pars$Bspe = 4 # Constant minimal number of interaction per species (extablishment prob)
pars$Bext = 4 # Constant minimal number of interaction per species (extinction prob)
pars$I_max = 40 # Maximal number of interactioning species
```



```{r}

source("../code/functions_sim_comp.fac.R")

```



### set parameters simulation

```{r}
nsteps = 350 # Set the maximum timestep per simulation
nsim = 2 # Set the number of simulations

# Vector to record number of simulation needed to reach 100 valid simulations
needed_sim_pos <- c(rep(NA, nsim))
needed_sim_neg <- c(rep(NA, nsim))
```

### Seeds record matrix

```{r}
# List of used seed per simulation batch
reccorded_seed_pos_list <- list()
reccorded_seed_neg_list <- list()

seed_list <- NULL # Set an object to record seed that will be used to launch simulations

```

## Simulation

```{r}
#for(b in 1:10){

	total_nb_sim <- 0
	b <- 1

	for(int in 0:1){
		pars$int = int # Define the interaction type: 0 = competition ; 1 = facilitation
		successful_sim <- 1 # Set the count of simulations
		list_res <- list() # Set the list to store the results

		if(pars$int == 1){
			print("Positive interctions")
		} else {
			print("Negative interctions")
		}

		seed_record <- c(rep(NA, nsim)) # Set the vector to record the seed that have been used

		while(successful_sim < nsim+1){

  		seed <- sample.int(80000, 1) # Pick a random number, this number will be the seed for one simulation
  		print(seed)

			# If a seed have already been tested, pick another one
			if(length(which(seed_list == seed)) != 0){
				while(seed %in% seed_list){
    			seed <- sample.int(80000, 1)
  			}
			}

			seed_list <- c(seed_list, seed) # Add the seed

			# SIMULATION ; from the file code/functions/functions_M-C_bifurcation.R
  		simulation <- sim_model_bif(seed = seed, pars, nsteps = nsteps)  
  	
  		#---------------------- stored: pres, traits_mat, anc, extinct, Stotact, L, L_list

			total_nb_sim <- total_nb_sim +1 # Count the total amount of simulations

			# Test if we have enough species at the timestep 150 in the simulation
  		success <- sum(simulation$pres[150,])
			#print(paste0("Success = ", success))

  		if(success >= 20){
				#print("the simulation has more than 19 species at the time step 150")
    		seed_record[successful_sim] <- seed # Record seed which alows us to have "good" simulation
    		res_sim <- list(simulation_number = paste0("simulation", successful_sim), seed = seed, parameters = pars, presence_matrix = simulation$pres, traits_df = simulation$traits, parentage_matrix = simulation$anc, extinxction_matrix = simulation$extinct, network_list = simulation$L_list, anc_dist_table = simulation$dist_anc, list_anc_dist = simulation$list_dist_anc) # Record results from the simulation
    		list_res <- list_append(list_res, res_sim)

    		print(paste("simulation ", successful_sim, "sur", nsim))
    		successful_sim <- successful_sim + 1 # Count the amount of "good" simulations
  		}
		}

		## Record the amount of simulation needed and their identity
		if(pars$int == 1){
			reccorded_seed_pos_list <- list_append(reccorded_seed_pos_list, seed_record)
			needed_sim_pos[b] <- total_nb_sim
		} else {
			reccorded_seed_neg_list <- list_append(reccorded_seed_neg_list, seed_record)
			needed_sim_neg[b] <- total_nb_sim
		}

		# Save simulations results data
		if(pars$int == 1){
			path <- paste0("../data/positive_interactions/list_res_pos_", nsim, ".Rdata")
	  	save(list_res, file = path) #Change if interaction tyep = 0 (competition)
		} else if(pars$int == 0){
			path <- paste0("../data/negative_interactions/list_res_neg_", nsim, ".Rdata")
	  	save(list_res, file = path) #Change if interaction tyep = 0 (competition)
		}

	}

	#print(paste("batch ", b, "sur 10"))
#}
	
```



### save data

```{r}
save(reccorded_seed_pos_list, file = "../AF_Data/positive_interactions/reccorded_seed_pos_list1.Rdata")
save(reccorded_seed_neg_list, file = "../AF_Data/negative_interactions/reccorded_seed_neg_list1.Rdata")
save(needed_sim_pos, file = "../Data/positive_interactions/needed_sim_pos1.Rdata")
save(needed_sim_neg, file = "../Data/negative_interactions/needed_sim_neg1.Rdata")

```


--- How results are stored
They are stored in list_res_pos_nsim.Rdata and list_res_neg_nsim.Rdata in the form of a list (list_res), which is a list of simulations, and each of them contains a list of the results of the simulation (res_sim).
I need to think about a way to combine the results of the simulations (phylog. signal) -> maybe statistic describing the relationship between phylog. signal and time, dimensionality, etc..

--- The simulation ends if it reaches either a maximum number of timesteps (350) or a maximum number of species (1000).
To make sure the list of networks and the list of trees per timesteps have (1) the same length and (2) are correspondent with the timesteps, I need to make sure that both start and finish to be measured at the same timesteps, and that I know then number of timesteps run per simulation. 


Start: 5 species present could be a good number for starting to measure both networks and phylogenies. 
For the trees, I calculate them from the anc table.
Networks are stored in a list.

So list_networks[[5]] should be compared with anc[[timestep 5]]


--- Phylogenetic distances
Record them myself rather than with hclust or any other algorythm.
Pass distances data to Netweek format.




# FOODWEB

### set general parameters

```{r}
rm(list = ls())

pars = list()

# General parameters
#pars$int = 0 # 0 for competition, 1 for facilitation, 2 for predation
pars$Smax = 1000 # Maximal number of species in the system

# Related to interaction range trait
pars$av_r = 0.1 # Half range of the niche of the first species
pars$sd = 0.5*pars$av_r + 0.0001 # Standard deviation of the normal distribution used to calculate the niche optimum trait of a new species

pars$u_max = 0.15 # Speciation probability

# Extinction probability, negative interactions communities
pars$a_eneg = 0.025 # Shape of the exponential decay of the negative extinction - interaction relationship
pars$e_0neg = 0.5 # Asymptotic extinction probability with infinite negative interactions
pars$e_1neg = 1  # Extinction probability with absence of interactions

# Extinction probability, positive interactions communities
pars$a_epos = 1.2  # Shape of the exponential decay of the positive extinction - interaction relationship
pars$e_0pos = 0.075 # Asymptotic extinction probability with infinite positive interactions
pars$e_1pos = 5.19 # 1 - pars$e_0pos

# Establishment probability, negative interactions communities
pars$a_uneg = 0.075 # Shape of the exponential decay of the colonization - interaction relationship
pars$u_0neg = 0.075 # Asymptotic establishment probability with infinite competition interactions
pars$u_1neg = 2 # Establishment probability with absence of competition interaction

# Establishment probability, positive interactions communities
pars$a_u = 0.45 # Shape of the exponential decay of the colonization - interaction relationship
pars$u_0 = 1 # Asymptotic establishment probability with infinite facilitation interactions
pars$u_1 = -1 # Establishment probability with absence of facilitation interactions
pars$d = 0.5 # Decrease speed of the establishment probability

# Extinction and establishment probabilitie for both, positive and negative interactions communities
pars$Bspe = 4 # Constant minimal number of interaction per species (extablishment prob)
pars$Bext = 4 # Constant minimal number of interaction per species (extinction prob)
pars$I_max = 40 # Maximal number of interactioning species
```


```{r}
pars$e_0neg = 0.1 # Asymptotic extinction probability with infinite negative interactions
```




```{r}

source("./functions_sim_foodweb.R")

```



### set parameters simulation

```{r}
nsteps = 350 # Set the maximum timestep per simulation
nsim = 1 # Set the number of simulations

# Vector to record number of simulation needed to reach 100 valid simulations
needed_sim_fw <- c(rep(NA, nsim))
```

### Seeds record matrix

```{r}
# List of used seed per simulation batch
reccorded_seed_fw_list <- list()

seed_list <- NULL # Set an object to record seed that will be used to launch simulations

```


```{r}

#for(b in 1:10){

total_nb_sim <- 0
b <- 1

successful_sim <- 1 # Set the count of simulations
list_res <- list() # Set the list to store the results


seed_record <- c(rep(NA, nsim)) # Set the vector to record the seed that have been used




while(successful_sim < nsim+1){
  
  seed <- sample.int(80000, 1) # Pick a random number, this number will be the seed for one simulation
  print(seed)
  
  # If a seed have already been tested, pick another one
  if(length(which(seed_list == seed)) != 0){
    while(seed %in% seed_list){
      seed <- sample.int(80000, 1)
    }
  }
  
  seed_list <- c(seed_list, seed) # Add the seed
  
  # SIMULATION ; from the file code/functions/functions_M-C_bifurcation.R
  simulation <- sim_model_bif(seed = seed, pars, nsteps = nsteps)  
  
  #---------------------- stored: pres, traits_mat, anc, extinct, Stotact, L, L_list
  
  total_nb_sim <- total_nb_sim +1 # Count the total amount of simulations
  
  # Test if we have enough species at the timestep 150 in the simulation
  success <- sum(simulation$pres[150,])
  #print(paste0("Success = ", success))
  
  if(success >= 1){
    #print("the simulation has more than 19 species at the time step 150")
    seed_record[successful_sim] <- seed # Record seed which alows us to have "good" simulation
    res_sim <- list(simulation_number = paste0("simulation", successful_sim), 
                    seed = seed, 
                    parameters = pars, 
                    presence_matrix = simulation$pres, 
                    traits_df = simulation$traits, 
                    parentage_matrix = simulation$anc, 
                    extinxction_matrix = simulation$extinct, 
                    network_list = simulation$L_list, 
                    anc_dist_table = simulation$dist_anc, 
                    list_anc_dist = simulation$list_dist_anc) # Record results from the simulation
    
    list_res <- list_append(list_res, res_sim)
    
    print(paste("simulation ", successful_sim, "sur", nsim))
    successful_sim <- successful_sim + 1 # Count the amount of "good" simulations
  }
  
}



  reccorded_seed_fw_list <- list_append(reccorded_seed_fw_list, seed_record)
  needed_sim_fw[b] <- total_nb_sim




#print(paste("batch ", b, "sur 10"))
#}
  
  path <- paste0("../data/foodweb/list_res_fw_", nsim, ".Rdata")
	  	save(list_res, file = path) 
  
```


### save data

```{r}

save(reccorded_seed_fw_list, file = "../data/foodweb/reccorded_seed_fw_list1.Rdata")
save(needed_sim_fw, file = "../data/foodweb/needed_sim_fw.Rdata")

```








# NEUTRAL 


# FOODWEB

### set general parameters

```{r}
rm(list = ls())

pars = list()

# General parameters
#pars$int = 0 # 0 for competition, 1 for facilitation, 2 for predation
pars$Smax = 1000 # Maximal number of species in the system

# Related to interaction range trait
pars$av_r = 0.1 # Half range of the niche of the first species
pars$sd = 0.5*pars$av_r + 0.0001 # Standard deviation of the normal distribution used to calculate the niche optimum trait of a new species

pars$u_max = 0.15 # Speciation probability

# Extinction probability, negative interactions communities
pars$a_eneg = 0.025 # Shape of the exponential decay of the negative extinction - interaction relationship
pars$e_0neg = 0.5 # Asymptotic extinction probability with infinite negative interactions
pars$e_1neg = 1  # Extinction probability with absence of interactions

# Extinction probability, positive interactions communities
pars$a_epos = 1.2  # Shape of the exponential decay of the positive extinction - interaction relationship
pars$e_0pos = 0.075 # Asymptotic extinction probability with infinite positive interactions
pars$e_1pos = 5.19 # 1 - pars$e_0pos

# Establishment probability, negative interactions communities
pars$a_uneg = 0.075 # Shape of the exponential decay of the colonization - interaction relationship
pars$u_0neg = 0.075 # Asymptotic establishment probability with infinite competition interactions
pars$u_1neg = 2 # Establishment probability with absence of competition interaction

# Establishment probability, positive interactions communities
pars$a_u = 0.45 # Shape of the exponential decay of the colonization - interaction relationship
pars$u_0 = 1 # Asymptotic establishment probability with infinite facilitation interactions
pars$u_1 = -1 # Establishment probability with absence of facilitation interactions
pars$d = 0.5 # Decrease speed of the establishment probability

# Extinction and establishment probabilitie for both, positive and negative interactions communities
pars$Bspe = 4 # Constant minimal number of interaction per species (extablishment prob)
pars$Bext = 4 # Constant minimal number of interaction per species (extinction prob)
pars$I_max = 40 # Maximal number of interactioning species

```



```{r}

p_est <- c(0.8, 0.9, 0.95)

p_ext <- c(0.001, 0.01, 0.1, 0.5)

df_scenarios <- expand.grid(p_est, p_ext)

colnames(df_scenarios) <- c("p_est", "p_ext")


```




```{r}

source("./functions_sim_neutral.R")

```



### set parameters simulation

```{r}
nsteps = 350 # Set the maximum timestep per simulation
nsim = 1 # Set the number of simulations

# Vector to record number of simulation needed to reach 100 valid simulations
needed_sim_neutral <- c(rep(NA, nsim))
```

### Seeds record matrix

```{r}
# List of used seed per simulation batch
reccorded_seed_neutral_list <- list()

seed_list <- NULL # Set an object to record seed that will be used to launch simulations

```


```{r}

sc_success <- numeric(nrow(df_scenarios))
fail_sim <- 1
list_objects_sc <- list()




for (sc in 1:nrow(df_scenarios)) {
 
  
  
  #for(b in 1:10){

total_nb_sim <- 0
b <- 1

successful_sim <- 1 # Set the count of simulations
list_res <- list() # Set the list to store the results


seed_record <- c(rep(NA, nsim)) # Set the vector to record the seed that have been used




while(successful_sim < nsim+1 & fail_sim != 40){
  
  seed <- sample.int(80000, 1) # Pick a random number, this number will be the seed for one simulation
  print(seed)
  
  # If a seed have already been tested, pick another one
  if(length(which(seed_list == seed)) != 0){
    while(seed %in% seed_list){
      seed <- sample.int(80000, 1)
    }
  }
  
  seed_list <- c(seed_list, seed) # Add the seed
  
  # SIMULATION ; from the file code/functions/functions_M-C_bifurcation.R
  simulation <- sim_model_bif(seed = seed, pars, nsteps = nsteps, 
                              pes = df_scenarios[sc, "p_est"],
                              pex = df_scenarios[sc, "p_ext"])  
  
  #---------------------- stored: pres, traits_mat, anc, extinct, Stotact, L, L_list
  
  total_nb_sim <- total_nb_sim +1 # Count the total amount of simulations
  
  # Test if we have enough species at the timestep 150 in the simulation
  success <- sum(simulation$pres[50,])
  #print(paste0("Success = ", success))
  
  if(success >= 5){
    #print("the simulation has more than 19 species at the time step 150")
    seed_record[successful_sim] <- seed # Record seed which alows us to have "good" simulation
    res_sim <- list(simulation_number = paste0("simulation", successful_sim), 
                    seed = seed, 
                    parameters = pars, 
                    presence_matrix = simulation$pres, 
                    traits_df = simulation$traits, 
                    parentage_matrix = simulation$anc, 
                    extinxction_matrix = simulation$extinct, 
                    network_list = simulation$L_list, 
                    anc_dist_table = simulation$dist_anc, 
                    list_anc_dist = simulation$list_dist_anc) # Record results from the simulation
    
    list_res <- list_append(list_res, res_sim)
    
    print(paste("simulation ", successful_sim, "sur", nsim))
    successful_sim <- successful_sim + 1 # Count the amount of "good" simulations
    
    sc_success[sc] <- "success"
    
  } else {
    
    sc_success[sc] <- "fail"
    
    fail_sim <- fail_sim + 1
    
  }
  
  
  
  
}



reccorded_seed_neutral_list <- list_append(reccorded_seed_neutral_list, seed_record)
needed_sim_neutral[b] <- total_nb_sim

list_objects_sc[[sc]] <- c(list_res, reccorded_seed_neutral_list)


#print(paste("batch ", b, "sur 10"))
#}
  
  
  
  
   
}


  
  path <- paste0("../data/neutral/list_res_neutral_", nsim, ".Rdata")
	  	save(list_objects_sc, file = path) 
  
```


### save data

```{r}

save(reccorded_seed_neutral_list, file = "../data/neutral/reccorded_seed_neutral_list1.Rdata")
save(needed_sim_neutral, file = "../data/neutral/needed_sim_neutral.Rdata")

```




Quick inspection of spp richness


```{r}

load("../data/neutral/list_res_neutral_1.Rdata")

list_results <- list_objects_sc[[1]]

results <- list_results[[1]]





n_spp_time <- c()

for (i in 1:nrow(results$presence_matrix)) {
  
  n_spp_time[i] <- length(which(results$presence_matrix[i,] == 1))
  
}

timesteps <- 1:nrow(results$presence_matrix)

df_signal_time<- as.data.frame(n_spp_time,timesteps)


library(ggplot2)

p.spp_neg <- ggplot(df_signal_time, aes(x=timesteps, y = n_spp_time)) +
  geom_line(color="black", linetype="twodash") +
  theme_bw()+
  #my.theme+
  scale_colour_manual(values = c("black", "red"))+
  ggtitle("Species richness through time (-)")+
  xlab("timesteps")+
  ylab("N species")

p.spp_neg

```


