---
title: "Evaluation of simulation FOODWEB - testing parameters "
output: html_notebook
description: this script runs one or few foodweb simulations and compares the characteristics of the simulated community with those of empirical foodwebs - different combinations of parameters can be tested to explore the results they yield
---





```{r}

library(igraph)
library(reshape2)
library(ggplot2)
library(ggpubr)


source("../../code/functions/functions_simulation_foodweb.R")
source("../../code/functions/functions_metrics_networks.R")
source("../../code/functions/functions_phylosignal.R")


```



For plotting:

```{r}
my.theme<-theme(axis.text=element_text(size=12),
                axis.title = element_text(size = 14),
                legend.position = "top",
                legend.text=element_text(size=10),
                legend.title = element_text(size=12),
                plot.title = element_text(face="bold",size=14,margin=margin(0,0,20,0),hjust = 0.5),
                axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)),
                axis.title.x = element_text(margin = margin(t = 15, r = 0, b = 0, l = 0)))
```




```{r}

plot_spp_richness <- function(steps, pres_mat, title) {
  
  timesteps <- 1:steps
  
  n_spp_time <- c()
  
  for (i in 1:nrow(pres_mat)) {
  
  n_spp_time[i] <- length(which(pres_mat[i,] == 1))
  
  }
  
  
  df_spprich_time <- data.frame(timesteps,n_spp_time)
  
  p.spp <- ggplot(df_spprich_time, aes(x=timesteps, y = n_spp_time)) +
  geom_line(color="black", linetype="twodash") +
  theme_bw()+
  my.theme+
  xlab("timesteps")+
  ylab("N species")+
    ggtitle(title)
  
  return(p.spp)
}




```





### Read table metrics from empirical foodwebs

empirical networks data was obtained from the ECOWeB 1.1 database:
https://digitalcommons.rockefeller.edu/cohen_joel_laboratory/1/

```{r}

net_metr_empirical <- readRDS("../../data/net_metrics_empirical.rds")

net_metr_empirical$id <- NULL

# eliminate one problematic foodweb:

net_metr_empirical <- net_metr_empirical[-13,]

# reshape dataframe:

df_metrics_emp <- reshape_df_net.metrics(df_metrics = net_metr_empirical,
                       interaction = "empirical")
```



# FOODWEB simulation


# set general parameters


```{r}

pars = list()
pars$u_max = 0.075 # mutation probability
pars$u_0 = 1  # 1 for predation & mutualism, 0 for competition
pars$u_1 = -1 # -1 for predation & mutualism, 1 for competition
pars$e_0neg = 0.1 # asymptotic extinction probability with infinite negative interactions
pars$e_1neg = -pars$e_0neg  # Extinction probability with absence of interactions
pars$e_0pos = 0.01 # asymptotic extinction probability with infinite positive interactions
pars$e_1pos = 1 - pars$e_0pos # 1 - e0
pars$a_u = 0.5 # shape of the exponential decay of the colonization - interaction relationship	
pars$a_eneg = 0.025 # shape of the exponential decay of the negative extinction - interaction relationship 
pars$a_epos = 0.5  # shape of the exponential decay of the positive extinction - interaction relationship 
pars$av_r = 0.272 # range of the niche 
pars$beta_n = 1 # parameter of the beta distribution 

pars$Sbasal = 25 # number of basal species 
pars$Smax = 1000 # Maximal number of species allowed

pars$d = 0.5 # Decrease speed of the establishment probability
pars$I_max = 60 # Maximal number of interactioning species


```


# parameters simulation

```{r}
nsteps =  250
n_sims = 3
list_adj_matrix.f <- list()

```


# Simulation

```{r}


list_results <- list()

for (i in 1:n_sims) {
  
  
  seed <- sample.int(80000, 1)
  
  results <- sim_model_bif(seed = seed, pars, nsteps = nsteps)
  
  names(results)[1] <- "presence_matrix"
  names(results)[7] <- "network_list"
  names(results)[9] <- "list_anc_dist"
  
  list_results[[i]] <- results
  
     
  list_adj_matrix.f[[i]] <- get_adjancency_from_simulation.fw_single(list_networks = results$network_list,
                                           nbasals = length(results$basal),
                                           pres_mat = results$presence_matrix)
  
}


# calculate metrics for each simulation

net_metr_sim <- get_network_measures_from_adjacency(list_adj_matrix = list_adj_matrix.f)

# compute table with mean metrics

df_metrics_fw <- reshape_df_net.metrics(df_metrics = net_metr_sim,
                       interaction = "foodweb")


# join with table of empirical networks

# Joining the empirical networks here

df_metrics <- merge(df_metrics_fw,
                    df_metrics_emp,
                    by = "metric")


# plot the results

p_S <- plot_metric_foodweb(df = df_metrics, metric = "S", y_axis = "N species")
p_link.dens <- plot_metric_foodweb(df = df_metrics, metric = "Link_density", y_axis = "L/S")
p_C <- plot_metric_foodweb(df = df_metrics, metric = "C", y_axis = "Connectance")
p_top <- plot_metric_foodweb(df = df_metrics, metric = "perc_tops", y_axis = "% top species")
p_int <- plot_metric_foodweb(df = df_metrics, metric = "perc_int", y_axis = "% intermediate species")
p_basal <- plot_metric_foodweb(df = df_metrics, metric = "perc_basals", y_axis = "% basal species")
p_cannibals <- plot_metric_foodweb(df = df_metrics, metric = "perc_cannibals", y_axis = "% cannibal species")
p_omnivory <- plot_metric_foodweb(df = df_metrics, metric = "omnivory", y_axis = "omnivory")
p_sd_gen <- plot_metric_foodweb(df = df_metrics, metric = "sd_gen", y_axis = "SD generality")
p_sd_vul <- plot_metric_foodweb(df = df_metrics, metric = "sd_vul", y_axis = "SD vulnerability")

plot_metrics_comparison_sim_emp <- ggarrange(
  p_S,
  p_link.dens,
  p_C,
  p_top,
  p_int,
  p_basal,
  p_cannibals,
  p_omnivory,
  p_sd_gen,
  p_sd_vul,
  
  nrow = 5,
  ncol = 2
  
)

plot_metrics_comparison_sim_emp

#ggsave("../../figures/simulation_evaluation/fw_metrics.png", height = 13, width = 9)

```




Pick simulation to inspect

```{r}
simulation1 <- list_results[[3]]
```


### Traits evolution

```{r, echo=FALSE}


pres = simulation1$presence_matrix
traits = simulation1$traits
nsteps = length(simulation1$network_list)
Time = c(1:nrow(pres))



#png(filename="../../figures/simulation_evaluation/traits_evol_fw.png", width = 700, height = 500, units = "px")


plot(Time[pres[,1]==1],rep(traits[1,1],sum(pres[,1])),xlim = c(1,nsteps),
ylim = c(0,1), pch = 19, cex = 0.3, xlab = "Time", ylab = "Niche position", main = "foodweb",
cex.axis = 2.25, cex.lab = 2.5)
for(i in 1:pars$Smax) points(Time[pres[,i]==1],rep(traits[i,1],sum(pres[,i])),cex = 0.3,pch = 19)


#dev.off()

```


# Degree distribution - Simulation

```{r}

for (i in 1:length(list_adj_matrix.f)) {
  
  graph <- graph_from_adjacency_matrix(adjmatrix = list_adj_matrix.f[[i]],
                                       mode = "directed")
  
  degree <- degree(graph, mode="all")
  
  deg.dist <- degree_distribution(graph, cumulative=T, mode="all")
  
 plot( x=0:max(degree), y=1-deg.dist, pch=19, cex=1.2,
                          xlab="Degree", ylab="Cumulative Frequency")
  
}


#png(filename="../../figures/simulation_evaluation/degree_distrib_fw.png", width = 700, height = 500, units = "px")


  graph <- graph_from_adjacency_matrix(adjmatrix = list_adj_matrix.f[[3]],
                                       mode = "directed")
  
  degree <- degree(graph, mode="all")
  
  deg.dist <- degree_distribution(graph, cumulative=T, mode="all")
  
 plot( x=0:max(degree), y=1-deg.dist, pch=19, cex=1.2,
                          xlab="Degree", ylab="Cumulative Frequency", main = "foodweb")
 
# dev.off()


```




### Speciation and extinction events

```{r, echo=FALSE}


# Number of speciation and extinction events
t0 = pres[1:(nsteps-1),]
t1 = pres[2:nsteps,] 
spec_mat = pres[1:(nsteps-1),]*0
ext_mat = pres[1:(nsteps-1),]*0
spec_mat[t1-t0==1] = 1 
ext_mat[t1-t0==-1] = 1
spec = apply(spec_mat ,1,sum)
ext = apply(ext_mat, 1, sum)

S = apply(pres,1,sum)[2:nsteps]




#png(filename="../../figures/simulation_evaluation/spec_ext_fw.png", width = 700, height = 500, units = "px")


plot(Time[2:nsteps], spec/S, type = "l", ylim = range(spec/S,ext/S),xlab = "Time", ylab = "Rate", col = "darkblue", main = "foodweb")
lines(Time[2:nsteps], ext/S, col = "darkred")
legend("topleft",bty = "n", legend = c("Speciation", "Extinction"), lty = 1, col = c("darkblue","darkred"))


#dev.off()

```

### Diversification-richness dependence

```{r, echo=FALSE}

#png(filename="../../figures/simulation_evaluation/div_rich_fw.png", width = 1300, height = 500, units = "px")



plot(S[2:nsteps],spec/S, xlab = "Species richness", ylab = "Rate per species", main = "foodweb", pch = 21, bg = "darkblue",cex = 0.8, cex.axis = 1.25, cex.lab = 1.5)
points(S[2:nsteps],ext/S, pch = 21, bg = "darkred",cex = 0.8)
legend("topright",bty = "n", legend = c("Speciation", "Extinction"), pch = 21, pt.bg = c("darkblue","darkred"))


#dev.off()

```




# plot species richness over time

```{r}

p_fw_rich <- plot_spp_richness(steps = nsteps, pres_mat = results$presence_matrix, title = "foodweb")

p_fw_rich

#ggsave("../../figures/simulation_evaluation/fw_richness.png", height = 6, width = 8)

```





# Degree distribution - Empirical networks


```{r}

# Make a vector of all your file paths

file_paths_n <- list.files(path = here::here("data/neutral/"), pattern = ".rds", full.names = TRUE)

# Get adjacency matrix from each simulation object stored in our folder
         
adj_matrix_list_n <-  get_adjancency_from_simulation.objects(file_path = file_paths_n)


# Compute network metrics

net_metr_n.sim <- get_network_measures_from_adjacency(list_adj_matrix = adj_matrix_list_n)

net_metr_n.sim$simulation <- NULL



# degree distribution

for (i in 1:length(adj_matrix_list_n)) {
  
  graph <- graph_from_adjacency_matrix(adjmatrix = adj_matrix_list_n[[i]],
                                       mode = "directed")
  
  degree <- degree(graph, mode="all")
  
  deg.dist <- degree_distribution(graph, cumulative=T, mode="all")
  
 plot( x=0:max(degree), y=1-deg.dist, pch=19, cex=1.2,
                          xlab="Degree", ylab="Cumulative Frequency")
  
}

    
```



